Evolution of Methodologies in Delivering Reliable Software in the European Scientific Arena

### 2018-06-11 paper-ieee-trans.tex
applied changes in introduction; added two references (aiftimiei and soltesz)

\IEEEPARstart{D}{uring} the last 15 years the provision of an European e--Infrastructure,
supporting unified access to large--scale computing and intense data analysis
applications, has been driven by the needs of scientific communities as part of their
pan--European research collaborations. One such example is the Worldwide
Large Hadron Collider (LHC) Computing Grid infrastructure, driven
by the particle physics scientific community, built to satisfy the computing needs
of processing, storing and analysis of the very large amount of data produced by the
LHC particle accelerator. A significant number of innovation projects and technology transfer
activities -- later discussed in Section~\ref{sec:ev} -- from the European Commission
(EC)-funded 5th, 6th and 7th Framework Programmes, and recently from the Horizon 2020 programme
\cite{h2020} had as their main goal to develop software solutions that provide a seamless and
transparent access of researchers to distributed computing and data resources.

Major developments carried throughout the years where focused on providing new
functionalities, while in the early days, the adoption of software engineering methodologies had a lower priority.
The released software was less tested, resulting in a higher rate of
rollbacks and patches applied to production systems, compared to the later projects. Thereby, the stability and
% Comment
%P1 C2 L18: Claiming that stability and reliability was regularly compromised is quite a bold statement -- it needs at least some citations, and preferably some explanation of what forms this compromise took for a few prominent cases.
reliability of the e--Infrastructures was often compromised, affecting
negatively their exploitation by the scientific communities.
% Comments
%P1 C2 L24: Was this realisation imposed "from above" e.g. by coordinated EC/ERC oversight, or was it "emergent" from the demands of the independent scientific communities?
It became then apparent the need for a better balance between the addition of new features
and the reliability of the released software. Consequently, incoming projects progressively
devoted a substantial part of the software lifecycle to the improvement of the quality of the
software being delivered. At this point, evolving software engineering practices were considered
and gradually adopted to define and implement quality procedures, the organization of
cross--functional teams and deployment of pilot testbeds.

Analysing the EC--funded software projects over the last decade,
one can observe a continuous increase in the prominence and robustness of software
testing and validation procedures. The
evolution of the software engineering methodologies (creation of the Agile
Manifesto \cite{agile-manifesto}, and rise of DevOps practices \cite{zhu}) together
with the parallel advancements in the information and communication technology
field (such as virtualization, the enabling technology for Cloud computing and
operating--system--level virtualization known as containerization) motivated this trend. All of which
are complemented by the emergence of automation and event--response tools.

@@

@@

>>>>

\IEEEPARstart{D}{uring} the last 15 years scientific communities have worked together to provide an European e--Infrastructure
able to support unified access to large--scale computing and intense data analysis
applications.
One such example is the Worldwide
Large Hadron Collider (LHC) Computing Grid infrastructure, which is driven
by the particle physics scientific community, built to satisfy the computing needs
of processing, storing and analysis of the very large amount of data produced by the
LHC particle accelerator. 
In the European Commission
(EC)-funded 5th, 6th and 7th Framework Programmes and recently in the Horizon 2020 programme
\cite{h2020}, a significant number of innovation projects and technology transfer
activities have to develop software solutions that provide a seamless and
transparent access of researchers to distributed computing and data resources.

Throughout the years, major developments have been focused on providing new
functionalities, while in the early days the adoption of software engineering procedures have had a higher priority.
The released software was less tested, causing a higher rate of
rollbacks and patches applied to production systems, compared to the software developed and maintained in the current projects. Thereby, the stability and
% Comment
%P1 C2 L18: Claiming that stability and reliability was regularly compromised is quite a bold statement -- it needs at least some citations, and preferably some explanation of what forms this compromise took for a few prominent cases.
reliability of the e--Infrastructures was often compromised, affecting
negatively their exploitation by the scientific communities.
% Comments
%P1 C2 L24: Was this realisation imposed "from above" e.g. by coordinated EC/ERC oversight, or was it "emergent" from the demands of the independent scientific communities?
Therefore, it has become clear the need for a better balance between the addition of new features
and the reliability of the released software. Consequently, incoming projects progressively
devoted a substantial part of the software lifecycle to the improvement of the quality of the
software being delivered. At this point, evolving software engineering practices have been considered
so far and gradually adopted to define and implement quality procedures, the organization of
cross--functional teams and deployment of pilot testbeds.

Analysing the EC--funded software projects over the last decade,
one can observe a continuous increase in the prominence and robustness of software
testing and validation procedures \cite{aiftimiei}. The
evolution of the software engineering methodologies (e.g., the creation of the Agile
Manifesto \cite{agile-manifesto} and the rise of DevOps practices \cite{zhu}) together
with the parallel advancements in the information and communication technology
field (e.g, the enabling technology for Cloud computing and the
operating--system--level virtualization through containerization \cite{soltesz}) motivated this trend. All of which
are complemented by the emergence of automation and event--response tools.

@@

\bibitem{aiftimiei}
Aiftimiei, A., Dongiovanni, D., et al.,  
Improving the quality of EMI Releases by leveraging the
EMI Testing Infrastructure.
\emph{Journal of Physics: Conference Series},
vol. 396, 2012.

@@

\bibitem{soltesz}
Soltesz, S., P\"{o}tzl, H., et. al.,
Container-based Operating System Virtualization: A Scalable, High-performance Alternative to Hypervisors,
\emph{SIGOPS Oper. Syst. Rev.},
vol. 41, n. 3, pp. 275--287, 2007.
